Cultural Transmission Can Explain the Evolution of Cooperation

### 2020-07-30 ms/main.tex
fix typos and their corresponding errors in the results, together with Dor

Individuals interact according to a prisoner's dilemma.
Specifically, individuals interact in pairs; a cooperator suffers a fitness cost $0<c<1$, and its partner gains a fitness benefit $b$, where we assume $b>c>0$. \textbf{\autoref{table:prisoner_payoff}} shows the payoff matrix, i.e. the fitness of an individual with phenotype $\phi_1$ when interacting with a partner of phenotype $\phi_2$.

@@

We define the initial frequency as $\tilde{p}_0$ and the cost boundaries
\begin{equation}\begin{aligned}
\gamma_1 = \frac{b \alpha T_A - (T_A - T_B)}{1-T_B}, \quad
\gamma_2 = \frac{b \alpha T_B + (1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}\end{equation}
Applying eqs.~\ref{eq:vert_hori_equilibrium}, \ref{eq:vert_hori_global_condition1}, and \ref{eq:vert_hori_global_condition2} we can summarize the possible outcomes:
%\begin{enumerate} % p* version
%\item Fixation of cooperation,
%	if $T_A>T_B$ and $0<\tilde{p}^*$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0>\tilde{p}^*$, or
%	if $T_A<T_B$ and $1<\tilde{p}^*$.
%\item Global fixation of defection,
%	if $T_A>T_B$ and $\tilde{p}^*>1$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0<\tilde{p}^*$, or
%	if $T_A<T_B$ and $\tilde{p}^*<0$
%\item Co-existence of both phenotypes at $\tilde{p}^*$, 
%	if $T_A<T_B$ and $0<\tilde{p}^*<1$.
%\end{enumerate}
\begin{enumerate} % gamma version
\item Fixation of cooperation,
	if $T_A>T_B$ and $c<\gamma_1$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0>\tilde{p}^*$; or
	if $T_A<T_B$ and $c<\gamma_2$.
\item Fixation of defection,
	if $T_A>T_B$ and $\gamma_2<c$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0<\tilde{p}^*$; or
	if $T_A<T_B$ and $\gamma_1<c$.
\item Coexistence of both phenotypes at $\tilde{p}^*$, 
	if $T_A<T_B$ and $\gamma_2<c<\gamma_1$.

Much of the literature on evolution of cooperation focuses on conditions for cooperation to invade a population of defectors.
The next corollary deals with such a condition.
\\
\end{enumerate}

@@

\begin{corollary}[No assortment of transmission and cooperation]
When $\alpha=0$, then either
\textbf{(1)} cooperation can fix, but not increase from rarity, if there is horizontal bias for cooperation ($T_A>T_B$) and the cost is low enough ($c<(1+b)(T_A-T_B)/(1-T_B)$), or 
\textbf{(2)} cooperation can increase from rarity to a stable coexistence with defection, if there is horizontal bias for defection ($T_A<T_B$) and the cost is low enough ($c<(T_B-T_A)/(1-T_B)$).
\end{corollary}

Here, the third equilibrium is
\begin{equation} \label{eq:vert_hori_alpha0_equilibrium}
\tilde{p}^*(\alpha=0) = \frac{c(1-T_B) - (T_A-T_B)}{b (T_A-T_B) },
\end{equation} 
and the cost boundaries are
\begin{equation}\begin{aligned}
\gamma_1(\alpha=0) = \frac{T_B - T_A}{1-T_B}, \quad
\gamma_2(\alpha=0) = \frac{(1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}\end{equation}
If $T_A>T_B$ then $\gamma_1(\alpha=0)<0<\gamma_2(\alpha=0)$.
So cooperation cannot increase from rarity, but it can fix if $c<\gamma_2(\alpha=0)$.
If $T_A<T_B$ then $\gamma_2(\alpha=0)<0<\gamma_1(\alpha=0)$.
So cooperation cannot fix, but if $c<\gamma_1(\alpha=0)$ then it can increase from rarity to a stable coexistence at $\tilde{p}^*(\alpha=0)$.
%TODO add "effective cost" argument: c<TA-TB/(1-TB) <=> 1-(1-c)(1-TB)<TA.
\\

>>>>

Individuals interact according to a prisoner's dilemma.
Specifically, individuals interact in pairs; a cooperator suffers a fitness cost $0<c<1$, and its partner gains a fitness benefit $b$, where we assume $c<b$. \textbf{\autoref{table:prisoner_payoff}} shows the payoff matrix, i.e. the fitness of an individual with phenotype $\phi_1$ when interacting with a partner of phenotype $\phi_2$.

@@

We define the initial frequency as $\tilde{p}_0$ and the cost boundaries
\begin{equation}\begin{aligned}
\gamma_1 = \frac{b \alpha T_A + (T_A - T_B)}{1-T_B}, \quad
\gamma_2 = \frac{b \alpha T_B + (1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}\end{equation}
Applying eqs.~\ref{eq:vert_hori_equilibrium}, \ref{eq:vert_hori_global_condition1}, and \ref{eq:vert_hori_global_condition2} we can summarize the possible outcomes:
%\begin{enumerate} % p* version
%\item Fixation of cooperation,
%	if $T_A>T_B$ and $0<\tilde{p}^*$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0>\tilde{p}^*$, or
%	if $T_A<T_B$ and $1<\tilde{p}^*$.
%\item Global fixation of defection,
%	if $T_A>T_B$ and $\tilde{p}^*>1$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0<\tilde{p}^*$, or
%	if $T_A<T_B$ and $\tilde{p}^*<0$
%\item Co-existence of both phenotypes at $\tilde{p}^*$, 
%	if $T_A<T_B$ and $0<\tilde{p}^*<1$.
%\end{enumerate}
\begin{enumerate} % gamma version
\item Fixation of cooperation,
	if $T_A>T_B$ and $c<\gamma_1$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0>\tilde{p}^*$; or
	if $T_A<T_B$ and $c<\gamma_2$.
\item Fixation of defection,
	if $T_A>T_B$ and $\gamma_2<c$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0<\tilde{p}^*$; or
	if $T_A<T_B$ and $\gamma_1<c$.
\item Coexistence of both phenotypes at $\tilde{p}^*$, 
	if $T_A<T_B$ and $\gamma_2<c<\gamma_1$.
\end{enumerate}

Much of the literature on evolution of cooperation focuses on conditions for cooperation to invade a population of defectors.
The next corollary deals with such a condition, followed by a corollary that deals with symmetric horizontal transmission, i.e. $T_A=T_B$.
\\

@@

\begin{corollary}[No assortment of transmission and cooperation]
When $\alpha=0$, then if there is horizontal bias for cooperation ($T_A>T_B$) and
\textbf{(1)} the cost is low compared to the bias ($c<(T_A-T_B)/(1-T_B)$), then cooperation will fix from any positive frequency; or
\textbf{(2)} the cost is low compared to the benefit ($c<(1+b)(T_A-T_B)(1-T_B)$), then cooperation will fix if the initial frequency if high enough ($\tilde{p}_0 > \tilde{p}^*$).
\end{corollary}

Here, the third equilibrium is
\begin{equation} \label{eq:vert_hori_alpha0_equilibrium}
\tilde{p}^*(\alpha=0) = \frac{c(1-T_B) - (T_A-T_B)}{b (T_A-T_B) },
\end{equation} 
and the cost boundaries are
\begin{equation}\begin{aligned}
\gamma_1(\alpha=0) = \frac{T_A - T_B}{1-T_B}, \quad
\gamma_2(\alpha=0) = (1+b)\frac{T_A - T_B}{1-T_B}.
\end{aligned}\end{equation}
If $T_A>T_B$ then $0<\gamma_1(\alpha=0)<\gamma_2(\alpha=0)$.
So either $c<\gamma_1(\alpha=0)$ or $\gamma_1(\alpha=0)<c<\gamma_2(\alpha=0)$ will allow fixation of cooperation, the latter only if the initial frequency is high enough.
If $T_A<T_B$ then $\gamma_2(\alpha=0)<\gamma_1(\alpha=0)<0<c$.
So defection will fix..
%TODO add "effective cost" argument: c<TA-TB/(1-TB) <=> 1-(1-c)(1-TB)<TA.
\\
